Symbolic Abstract Data Type Inference

### 2015-10-25 histories.tex
Algorithm cleanup.
Fixed typos from Reviewer A.
Corrections from Reviewer B.

\end{itemize}
The \emph{frequency} of an operation $o$ is the number of operations identical
to $o$, denoted $\freq(o)$. We say that $o$ has \emph{duplicates} when
$\freq(o) > 1$. An operation $o \in \img(m)$ in the image of $m$ is a
\emph{match target}, and the inverse set $m^{-1}(o)$ of a target is a
\emph{match}. We assume $o \in m^{-1}(o)$ for every $o$. A operation $o$ is \emph{unmatched}
when $o$ is completed and $m(o)$ is undefined.

The \emph{width} of a history is the maximum number of its operations which are
mutually unordered, and the \emph{width} of a history set is the maximum width
of its elements. Width-$1$ histories are \emph{sequential}. The \emph{weight}
of a history is the maximum among operation frequencies and the number of
matches. Since the operation identifiers of a history have no intrinsic
meaning, we consider equality between histories up to renaming of operation
identifiers.

@@

\end{itemize}
We denote the set $\set{H(e,M,R) : e \in E}$ of histories of an execution set
$E$ by $H(E,M,R)$. When the matching scheme $\tup{M,E}$ is clear from the
context, we abbreviate $H(e,M,R)$ and $H(E,M,R)$ by $H(e)$ and $H(E)$.

>>>>

\end{itemize}
The \emph{frequency} of an operation $o$ is the number of operations identical
to $o$, denoted $\freq(o)$. We say that $o$ has \emph{duplicates} when
$\freq(o) > 1$. An operation $o \in \img(m)$ in the image of $m$ is a
\emph{match target}, and the inverse set $m^{-1}(o)$ of a target is a
\emph{match}. We assume $o \in m^{-1}(o)$ for every $o \in \img(m)$.
An operation $o$ is \emph{unmatched} when $o$ is completed and $m(o)$ is undefined.

The \emph{width} of a history is the maximum number of its operations which are
mutually unordered, and the \emph{width} of a history set is the maximum width
of its elements. Width-$1$ histories are \emph{sequential}. Since the operation
identifiers of a history have no intrinsic meaning, we consider equality between
histories up to renaming of operation identifiers.

@@

\end{itemize}
We denote the set $\set{H(e,M,R) : e \in E}$ of histories of an execution set
$E$ by $H(E,M,R)$. When the matching scheme $\tup{M,R}$ is clear from the
context, we abbreviate $H(e,M,R)$ and $H(E,M,R)$ by $H(e)$ and $H(E)$.

### 2015-10-25 inference.tex
Corrections from Reviewer B.

A Set $K$ \emph{generates} $H$ when the closure of $K$ under the relation
$\mathord\to = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c})$
is equal to $H$, i.e., $H = \set{ h : \exists h' \in K.\ h' \to^\ast h }$.
A \emph{kernel} of a set $H$ is a minimal set generating $H$. While the kernels
of arbitrary sets need not be unique, the kernels of sets which
have sequential kernels are unique. Furthermore, Section~\ref{sec:nature}
demonstrates that the histories of naturally-occurring implementations have
unique kernels, which we assume for the remainder of this work.
An \emph{abstract data type (ADT)} $A$ is the kernel of the set $H(\mathcal{I})$
of histories of some implementation $\mathcal{I}$; we say that \emph{$A$ is the
ADT of $\mathcal{I}$}.

>>>>

A Set $K$ \emph{generates} $H$ when the closure $K^*$ of $K$ under the relation
$\mathord\to = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c})$
is equal to $H$, i.e., $H = \set{ h : \exists h' \in K.\ h' \to^\ast h }$.
A \emph{kernel} of a set $H$ is a minimal set generating $H$. While the kernels
of arbitrary sets need not be unique, the kernels of sets which
have sequential kernels are unique. Furthermore, Section~\ref{sec:nature}
demonstrates that the histories of naturally-occurring implementations have
unique kernels, which we assume for the remainder of this work.
An \emph{abstract data type (ADT)} $A$ is the kernel of the set $H(\mathcal{I})$
of histories of some implementation $\mathcal{I}$; we say that \emph{$A$ is the
ADT of $\mathcal{I}$}.

### 2015-10-25 patterns.tex
Corrections from Reviewer B.

Besides this closure property, the $\preceq$ relation enjoys a certain notion of
well-foundedness when restricted to bounded-width histories: the set of
$\preceq$-minimal elements of any (potentially infinite) history set is finite.
This property is what enables us to represent infinite sets of invalid ADT
histories with a finite set of minimal examples. This property is captured
formally with wqos: a \emph{well-quasi-ordering (wqo)} $R$ on a set $X$ is a
reflexive, transitive binary relation on $X$ for which in every infinite
sequence $x_0 x_1 \ldots$ of elements from $X$, there exists indices $i < j$
such that $R(x_i,x_j)$.

@@

For the remainder of this section, we fix an ADT $A$, and let $B$ be its bounded
complement. When $A$ has bounded width, so does $B$, and thus $\preceq$ is a wqo
on $B$. Furthermore, when $A$ normal, it is closed under $\succeq$, and thus $B$
is closed\footnote{Actually, $B$ is closed under $\preceq$ when restricted to
width-bounded histories, i.e., if $h_1 \preceq h_2$, $h_1 \in B$, and $h_2$ is
width-bounded, then $h_2\in B$.} under $\preceq$. Closure under a relation
satisfying Lemma~\ref{lem:wqo} implies representation by a finite set. Formally,
we say a set $X$ is \emph{finitely representable} if there exists a finite set
$Y$ and a relation $R \subseteq Y \times X$ such that $X = \set{ x : \exists
y\in Y.\ R(y,x) }$. In our case, we obtain a finite set from which exactly the
elements of $B$ are related by $\preceq$.

>>>>

Besides this closure property, the inverse $\preceq$ relation enjoys a certain notion of
well-foundedness when restricted to bounded-width histories: the set of
$\preceq$-minimal elements of any (potentially infinite) history set is finite.
This property is what enables us to represent infinite sets of invalid ADT
histories with a finite set of minimal examples. This property is captured
formally with wqos: a \emph{well-quasi-ordering (wqo)} $R$ on a set $X$ is a
reflexive, transitive binary relation on $X$ for which in every infinite
sequence $x_0 x_1 \ldots$ of elements from $X$, there exists indices $i < j$
such that $R(x_i,x_j)$.

@@

For the remainder of this section, we fix an ADT $A$, and let $B$ be its bounded
complement. When $A$ has bounded width, so does $B$, and thus $\preceq$ is a wqo
on $B$. Furthermore, when $A$ is normal, it is closed under $\succeq$, and thus $B$
is closed\footnote{Actually, $B$ is closed under $\preceq$ when restricted to
width-bounded histories, i.e., if $h_1 \preceq h_2$, $h_1 \in B$, and $h_2$ is
width-bounded, then $h_2\in B$.} under $\preceq$. Closure under a relation
satisfying Lemma~\ref{lem:wqo} implies representation by a finite set. Formally,
we say a set $X$ is \emph{finitely representable} if there exists a finite set
$Y$ and a relation $R \subseteq Y \times X$ such that $X = \set{ x : \exists
y\in Y.\ R(y,x) }$. In our case, we obtain a finite set from which exactly the
elements of $B$ are related by $\preceq$.

### 2015-10-25 related.tex
Corrections from Reviewer B.

Inference of symbolic abstract data types is a recently-pertinent problem
arising from the the emergence of efficient symbolic refinement-checking
algorithms~\cite{conf/popl/BouajjaniEEH15, conf/pldi/EmmiEH15} which require
symbolic, logical representations of ADT specifications. Though approaches
based on the explicit enumeration of execution
linearizations~\cite{journals/jpdc/WingG93, conf/pldi/BurckhardtDMT10,
conf/asplos/BurnimNS11, conf/kbse/ZhangCW13} do not require symbolic ADT
representations, and work directly with given reference implementations, these
approaches are intractable as they elaborate an exponential\footnote{The number
of possible linarizations is exponential in execution length.} number of
linearizations.

>>>>

Inference of symbolic abstract data types is a recently-pertinent problem
arising from the emergence of efficient symbolic refinement-checking
algorithms~\cite{conf/popl/BouajjaniEEH15, conf/pldi/EmmiEH15} which require
symbolic, logical representations of ADT specifications. Though approaches
based on the explicit enumeration of execution
linearizations~\cite{journals/jpdc/WingG93, conf/pldi/BurckhardtDMT10,
conf/asplos/BurnimNS11, conf/kbse/ZhangCW13} do not require symbolic ADT
representations, and work directly with given reference implementations, these
approaches are intractable as they elaborate an exponential\footnote{The number
of possible linarizations is exponential in execution length.} number of
linearizations.
